Milk intake and stroke mortality in the Japan Collaborative Cohort Study - a Bayesian survival analysis

### 2020-08-25 Paperfolder/MDPI_submission/Revision01/Response/Response to the reviewers/Response to the reviewers.Rmd

The interaction effect by sex was found probably due to the fact that women in our cohort had relatively less mortality rates of fatal stroke and much better lifestyles (much less smokers and alcohol drinkers) than men, or maybe due to other factors that was not available in/considered by our models such as intake of calcium supplements. The existing beneficial effect by milk intake is probably less evident in women than that in men in the JACC study cohort.

The updated manuscript have included these information in the methods section as well as in the discussion about sex differences on the association.

3. *The aim states to “provide a more straightforward answer to the primary research question” – compared to what? And how is this more straightforward answer achieved. This is not clear, neither in the aim nor in the discussion (lines166-167). Please revise.*

[Response:] Thanks for your suggestion. The more straightforward answer is provided as Pr(HR < 1) which shows the percentages of HRs < 1 calculated through the large amount of simulations. The detailed interpretation is described above in answering your second comment. These probabilities shows directly how certain it is for individuals with that milk intake frequency could have lower hazard compared with those who never consume milk. In traditional way the answer was not clear since only estimates of HRs, 95% confidence intervals, and/or P-values were shown. None of these actually directly answer the question "how sure it is to have lower hazard?" We have added clearer descriptions about these research questions and interpretations.

3. *Please describe the context of milk intake in Japan. Is milk consumed as a beverage with food, added to coffee or other consumption patterns? Are other milk products than fresh non-fermented milk consumed and could be included in the “milk” exposure?*

[Response:]

2. *Please specify age at baseline in abstract and methods.*
2. *Outcome ascertainment: Information on cause of death was from death certificates. Is there a possibility of individual linkage of the FFQ responses to a cause of death registry or were each of the participants’ death certificates reviewed? How many died from other causes, how many moved from the study area (and were thus lost to follow-up) and how many were censored at the end of follow-up (alive). Has stroke mortality changed during follow-up? Is there a difference in time trends of stroke subtype mortality? In how many cases was stroke subtype not recorded? The authors should also discuss potential reasons for differences in the results for stroke subtypes.*
2. *I agree with the authors that presenting the results in the FFQ categories is a good option, rather than to calculate the amount of consumption. However, I wonder if there is some indication from the validation study as to how much is on average consumed in the higher categories of intake?*
2. *Why did you adjust for sleep duration, is that a factor that influences both milk intake and stroke mortality risk?*
2. *The manuscript has minor language errors throughout and would benefit from a language review.*

### 2020-08-25 Paperfolder/MDPI_submission/Revision01/Response/Response to the reviewers/Response to the reviewers.Rmd

## Comments from reviewer #2

*The manuscript by C Wang et al includes a cohort of 94385 men and women from Japan, free from self-reported history of stroke, cancer, myocardial infarction and other types of heart disease at baseline who answered an FFQ regarding milk consumption. Between baseline in 1988-90 and end of follow-up in 2009, 2675 deaths from stroke were recorded on death registers. The authors applied Bayesian survival analysis to calculate hazard ratios and acceleration factors. With a range of milk consumption from “never” to “almost daily”, a higher milk consumption was associated with lower hazard and speed from dying from total stroke among men but not among women.*

*I have the following comments and questions for the authors:*

- *The introduction and discussion should however be more clear about the fact that the study is being performed in the lower range of the milk intake distribution. The dose-response meta-analysis in reference 2 indicates that the association may be non-linear. It would be good if the authors in more detail compared their results with other studies using fatal stroke as outcome and perhaps also focus on those cohorts with a low milk intake. For example, please elaborate on references 6, 20-23 in the discussion. Some of the references do not investigate milk intake per se. In the discussion, please specify that reference 19 does not investigate stroke mortality but CVD and all-cause mortality.*

@@

[Response:] Thanks for pointing these out. We have updated the introduction as well as the discussion parts in order to better showing that these participants actually represent those in the lower range of milk intake. Discussion comparing our findings with previous studies conducted in East Asian population has been expanded in more details. The fact that reference 19 was looking into CVD and all-cause mortality has been clarified.

- *I am unfortunately not familiar with the Bayesian models used in this work. I am sure that many readers will be in a similar situation as me. I appreciate that these models can calculate hazard ratios without fulfilling the proportional hazards assumption. The authors should in more detail describe the benefits and limitations of these models in relation to the more commonly used Cox proportional hazards regression. Is a HR estimated from these models directly comparable with that from a Cox model? How is an acceleration factor interpreted? Do the models assume that a higher milk intake is associated with a lower stroke mortality, indicated also by the Pr(HR<1)? If so, how would that influence the validity of the method since you in the introduction write that there are reports of inverse associations, no association, and direct associations? In Table 2, both SD and 95% CrI are given. Is there a reason for specifying the SD?*

[Response:] Thanks for your suggestions and comments on this. Acceleration factor (AF) can be interpreted as the speed/velocity just like all individuals are driving their cars from the same start point (entry into the study setting) to their goals (stroke mortality) but different people can drive their own cars with different speeds. If the AF is smaller than 1 (reference group - never drinker), then it is considered that individual will end up at the goal with slower speed (or takes longer time). Actually these AF maybe easier to be interpreted compared with hazard ratios (HR) for readers without any training in epidemiology. Because the idea of HR includes hazard that is defined as a risk at a given point in time which is more confusing for many readers.

We understand that some readers may not be familiar with the alternative and relatively computationally expensive approach. The biggest features in the traditional Cox proportional hazard model or any other classical statistical methods are that they assume 1) there is only one true but unknown value of the parameter (in this context, the HR and AF) from the data at hand to estimate; 2) the experiment/study can be repeated, therefore the confidence intervals in that framework uses standard error to provide precision for the estimates that can be interpreted as "if we repeat the same experiment for many times, 95% of these confidence intervals will contain the true but unknown value of interest"; 3) continue on the repeatable assumption, so that the P-values are actually saying that "if we can repeat the same experiment for many times, the same or more extreme results can happen with this probability, if this probability is lower than 0.05 (commonly used as the cut-off value but without actual solid scientific support) it is considered as rarely happens or 'statistically significant'."

Bayesian ways of statistical analyses would disagree with all of the above assumptions because literally no one experiment/study can be repeated especially large scale cohort studies that take years and many participants/investigators to carry out. Bayesian statistics would provide a posterior distribution of the parameter that we want to estimates which means that the HR or AF is not fixed and unknown but with a distribution that has a mean and a standard deviation to describe its features, and the 95% credible intervals is directly showing us that we are 95% sure that the parameter of interest is distributed in this range. Furthermore, because of the large amount of simulations that we have conducted, the Pr(HR < 1) is showing the percentages that these HRs are smaller than 1. So a percentage of Pr(HR < 1) = 99% can be directly interpreted as "we are 99% sure that drinking milk at this frequency has lower hazard than those who never consume milk." The models in both classical and the Bayesian frameworks are the same, so the HRs from both frameworks can be interpreted in the same way. However, the estimating processes are different: one is based on the maximum likelihood estimation but the other is based on extensive Markov Chain Monte Carlo (MCMC) simulation. Therefore we would say that the only limitation for the Bayesian approach would be the computational cost but with the development of the modern computers we expect the Bayesian approaches would become the future default. The above discussion would be out of the scope of the current paper but we understand the confusions could still exist even if we add them into the manuscript. As requested, we added some of the discussion in the revised manuscript in the "strength and limitation" section.

- *Why were analyses performed stratified by sex? Since no pooled analysis is presented, the reason for a stratified analysis should be stated in the aims, unless this was a secondary analysis and then also the results from the total cohort should be presented. The authors should also discuss the observed sex differences in more detail. Do other studies show similar sex differences? What potential biological mechanisms may explain them?*

[Response:] It is reasonable to conduct the analysis separated in men and women since we expected there is a sex difference in the association based on our prior knowledge from the same cohort as well as the Singapore Chinese Health Study. Our expectation was further confirmed through a quick likelihood ratio test comparing Cox proportional hazard models with and without the interaction terms between sex and milk intake frequencies (P for interaction < 0.0001, not shown in the main text). Our stratified analyses also suggested the AFs and HRs might be in slightly different directions among 2 out of 4 frequencies (Table 2 and Table 3, results in the "total stroke" parts), we therefore believe it is legitimate to conduct analyses stratified by sex.

The interaction effect by sex was found probably due to the fact that women in our cohort had relatively less mortality rates of fatal stroke and much better lifestyles (much less smokers and alcohol drinkers) than men regardless of milk intake frequency, or maybe due to other factors that was not available in/considered by our models such as intake of calcium or other supplements. The existing beneficial effect by milk intake is probably less evident in women than that in men in the JACC study cohort.

@@

The updated manuscript have included these information in the methods section as well as in the discussion about sex differences on the association.

- *The aim states to “provide a more straightforward answer to the primary research question” – compared to what? And how is this more straightforward answer achieved. This is not clear, neither in the aim nor in the discussion (lines166-167). Please revise.*

[Response:] Thanks for your suggestion. The more straightforward answer is provided as Pr(HR < 1) which shows the percentages of HRs < 1 calculated through the large amount of simulations. The detailed interpretation is described above in answering your second comment. These probabilities show directly how certain it is for individuals with that milk intake frequency could have lower hazard compared with those who never consume milk. In the traditional way the answer was not clear since only estimates of HRs, 95% confidence intervals, and/or P-values were shown. None of these actually can directly answer the question "how sure it is to have lower hazard in one group compared with another?" We have added clearer descriptions about these research questions and interpretations.

- *Please describe the context of milk intake in Japan. Is milk consumed as a beverage with food, added to coffee or other consumption patterns? Are other milk products than fresh non-fermented milk consumed and could be included in the “milk” exposure?*

[Response:] Thanks for your suggestions.   In the baseline year of the JACC study (between 1988 and 1990), most "milk and dairy product" consumption (92.1%) was in the form of whole milk. We have added the context of milk intake in Japan for participants who were invited into the JACC study.

- *Please specify age at baseline in abstract and methods.*

[Response:] Age range are now added in the abstract and methods accordingly.

- *Outcome ascertainment: Information on cause of death was from death certificates. Is there a possibility of individual linkage of the FFQ responses to a cause of death registry or were each of the participants’ death certificates reviewed? How many died from other causes, how many moved from the study area (and were thus lost to follow-up) and how many were censored at the end of follow-up (alive). Has stroke mortality changed during follow-up? Is there a difference in time trends of stroke subtype mortality? In how many cases was stroke subtype not recorded? The authors should also discuss potential reasons for differences in the results for stroke subtypes.*

[Response:] We have added these details of information at the beginning of the results section. In the study area, investigators conducted systematic review of death certificates till the end of 2009. Date and cause of death were confirmed with the permission of the Director-General of the Prime Minister's Office.

- *I agree with the authors that presenting the results in the FFQ categories is a good option, rather than to calculate the amount of consumption. However, I wonder if there is some indication from the validation study as to how much is on average consumed in the higher categories of intake?*
- *Why did you adjust for sleep duration, is that a factor that influences both milk intake and stroke mortality risk?*
- *The manuscript has minor language errors throughout and would benefit from a language review.*

### 2020-08-25 Paperfolder/Paperfolder.Rmd
updated with 3 q left

  The aim was to further examine the relationship between milk intake and stroke mortality among the Japanese population. We used data from the Japan Collaborative Cohort (JACC) Study to estimate the posterior acceleration factors (AF) as well as the hazard ratios (HR) comparing individuals with different milk intake frequencies against those who never consumed milk at the study baseline. These estimations were computed through a series of Bayesian survival models that employed a Markov Chain Monte Carlo simulation process. 100000 posterior samples were generated separately through four independent chains after model convergency were confirmed. Posterior probabilites that daily milk consumers had lower hazard or delayed mortality from strokes compared to non-consumers was 99.0% and 78.0% for men and women, respectively. Accordingly, the estimated posterior means of AF and HR for daily milk consumers were 0.88 (95% Credible Interval, CrI: 0.81, 0.96) and 0.80 (95% CrI: 0.69, 0.93) for men and 0.97 (95% CrI: 0.88, 1.10) and 0.95 (95% CrI: 0.80, 1.17) for women. In conclusion, data from the JACC study has provided strong evidence that daily milk intake among Japanese men was associated with delayed and lower hazard of mortality from stroke especially cerebral infarction.

@@

Moreover, a more intuitive interpretation would be available if we were able to show the probabilities and how certain the existing data can provide evidence about whether drinking milk can delay or lower the hazard of dying from stroke. Accelerated failure time models (AFT) under Bayesian framework are convenient tools that would help avoid worrying about the assumption of proportional hazard normally required in a Cox proportional hazard model. AFT model is also helpful in estimating acceleration factors in addition to hazard ratios which can be interpreted as the speed for developing an event of interest. This speed parameter shows how faster/slower individuals in one exposure group might have an event compared to others among different exposure groups [@Wei1992;@ibrahim2014b]. AFT models

Our aim was to provide a more straightforward answer to the primary research question that whether someone answered he/she drank milk at the baseline of study had lower hazard of dying from stroke compared with his/her counterparts who said they never consumed milk. If the answer to the primary objective was yes, then the probabilities that individuals with different frequencies of milk intake may had lower hazard compared with those who never drank milk were calculated through a Markov Chain Monte Carlo (MCMC) simulation process. A Bayesian survival analysis method was applied on an existing database and through which, we also provided estimates about whether drinking milk could delay a stroke mortality event from happening after controlling for the other potential confounders.

@@

We used data from the Japan Collaborative Cohort (JACC) study, which was sponsored by the Ministry of Education, Sports, Science, and Technology of Japan. Sampling methods and details about the JACC study have been described extensively in the literature [@Ohno2001; @Tamakoshi2005; @Tamakoshi2013]. Participants of the JACC study completed self-administered questionnaires about their lifestyles, food intake (food frequency questionnaire, FFQ), and medical histories of cardiovascular disease or cancer. In the final follow-up of the JACC study, data from a total of 110585 individuals (46395 men and 64190 women) were successfully retained for the current analysis. We further excluded samples if they meet one of the following criteria: 1) with any disease history of stroke, cancer, myocardial infarction, ischemic heart disease, or other types heart disease (n = 6655, 2931 men and 3724 women); 2) did not answer the question regarding their milk consumption in the baseline FFQ survey (n = 9545, 3593 men and 5952 women). Finally, 94385 (39386 men and 54999 women) are left in the database. The study design and informed consent procedure were approved by the Ethics Review Committee of Hokkaido University School of Medicine.

@@

Frequency of milk intake during the preceding year of the baseline was assessed by FFQ from "never", "1-2 times/month", "1-2 times/week", "3-4 times/week", and "Almost daily". The exact amount of milk consumption was difficult to assess here. However, good reproducibility and validity were confirmed previously (Spearman rank correlation coefficient between milk intake frequency and weighed dietary record for 12 days was 0.65) [@Date2005]. From the validation study [@Date2005] of the FFQ, daily consumers of milk was found to have a median intake of 146 gram per day.

>>>>

  The aim was to further examine the relationship between milk intake and stroke mortality among the Japanese population. We used data from the Japan Collaborative Cohort (JACC) Study (total number of participants = 110585, age range: 40-79) to estimate the posterior acceleration factors (AF) as well as the hazard ratios (HR) comparing individuals with different milk intake frequencies against those who never consumed milk at the study baseline. These estimations were computed through a series of Bayesian survival models that employed a Markov Chain Monte Carlo simulation process. 100000 posterior samples were generated separately through four independent chains after model convergency were confirmed. Posterior probabilites that daily milk consumers had lower hazard or delayed mortality from strokes compared to non-consumers was 99.0% and 78.0% for men and women, respectively. Accordingly, the estimated posterior means of AF and HR for daily milk consumers were 0.88 (95% Credible Interval, CrI: 0.81, 0.96) and 0.80 (95% CrI: 0.69, 0.93) for men and 0.97 (95% CrI: 0.88, 1.10) and 0.95 (95% CrI: 0.80, 1.17) for women. In conclusion, data from the JACC study has provided strong evidence that daily milk intake among Japanese men was associated with delayed and lower hazard of mortality from stroke especially cerebral infarction.

@@

Moreover, a more intuitive interpretation would be available if we were able to show the probabilities and how certain the existing data can provide evidence about whether drinking milk can delay or lower the hazard of dying from stroke. This would require transformation of the research questions into "Do individuals with milk intake habit have lower hazard of dying from a stroke event compared with people who never consume milk?" Another way of asking the same research question would be "Can a event of dying from stroke be postponed by having a milk intake habit?" The later question is indeed directly comparing the length of times before events or the speed from the same departure point (entry of the study) to a fatal stroke event. Accelerated failure time models (AFT) under Bayesian framework are convenient tools that would help avoid worrying about the assumptions of repeatability of a large scale prospective study, and can calculate the exact proportion of posterior estimates (a hazard ratio) that is lower than 1. AFT model is helpful in estimating acceleration factors which can be interpreted as the speed for developing an event of interest. This speed parameter shows how faster/slower individuals in one exposure group might have an event compared to others among different exposure groups [@Wei1992;@ibrahim2014b].

Our aim was to provide a more straightforward answer to the primary research question that whether someone answered he/she drank milk at the baseline of study had lower hazard of dying from stroke compared with his/her counterparts who said they never consumed milk. If the answer to the primary objective was yes, then the probabilities that individuals with different frequencies of milk intake may had lower hazard compared with those who never drank milk were calculated through a Markov Chain Monte Carlo (MCMC) simulation process. A Bayesian survival analysis method was applied on an existing database and through which, we provided estimates about whether drinking milk could delay a stroke mortality event from happening after controlling for the other potential confounders.

@@

We used data from the Japan Collaborative Cohort (JACC) study, which was sponsored by the Ministry of Education, Sports, Science, and Technology of Japan. Sampling methods and details about the JACC study have been described extensively in the literature [@Ohno2001; @Tamakoshi2005; @Tamakoshi2013]. Participants of the JACC study completed self-administered questionnaires about their lifestyles, food intake (food frequency questionnaire, FFQ), and medical histories of cardiovascular disease or cancer. In the final follow-up of the JACC study, data from a total of 110585 individuals (46395 men and 64190 women) who aged between 40-79 at the baseline year were successfully retained for the current analysis. We further excluded samples if they meet one of the following criteria: 1) with any disease history of stroke, cancer, myocardial infarction, ischemic heart disease, or other types heart disease (n = 6655, 2931 men and 3724 women); 2) did not answer the question regarding their milk consumption in the baseline FFQ survey (n = 9545, 3593 men and 5952 women). Finally, 94385 (39386 men and 54999 women) are left in the database. The study design and informed consent procedure were approved by the Ethics Review Committee of Hokkaido University School of Medicine.

@@

Frequency of milk intake during the preceding year of the baseline was assessed by FFQ from "never", "1-2 times/month", "1-2 times/week", "3-4 times/week", and "Almost daily". The exact amount of milk consumption was difficult to assess here. However, good reproducibility and validity were confirmed previously (Spearman rank correlation coefficient between milk intake frequency and weighed dietary record for 12 days was 0.65) [@Date2005]. From the same validation study [@Date2005] on the FFQ, daily consumers of milk was found to have a median intake of 146 gram per day. In the baseline year of the JACC study (between 1988 and 1990), most "milk and dairy product" consumption (92.1%) was in the form of whole milk [@eigyogenjyo].

### 2020-08-25 Paperfolder/Paperfolder.tex
updated with 3 q left

Moreover, a more intuitive interpretation would be available if we were
able to show the probabilities and how certain the existing data can
provide evidence about whether drinking milk can delay or lower the
hazard of dying from stroke. Accelerated failure time models (AFT) under
Bayesian framework are convenient tools that would help avoid worrying
about the assumption of proportional hazard normally required in a Cox
proportional hazard model. AFT model is also helpful in estimating
acceleration factors in addition to hazard ratios which can be
interpreted as the speed for developing an event of interest. This speed
parameter shows how faster/slower individuals in one exposure group
might have an event compared to others among different exposure groups
\citep{Wei1992, ibrahim2014b}. AFT models

Our aim was to provide a more straightforward answer to the primary
research question that whether someone answered he/she drank milk at the
baseline of study had lower hazard of dying from stroke compared with
his/her counterparts who said they never consumed milk. If the answer to
the primary objective was yes, then the probabilities that individuals
with different frequencies of milk intake may had lower hazard compared
with those who never drank milk were calculated through a Markov Chain
Monte Carlo (MCMC) simulation process. A Bayesian survival analysis
method was applied on an existing database and through which, we also
provided estimates about whether drinking milk could delay a stroke
mortality event from happening after controlling for the other potential
confounders.

@@

Frequency of milk intake during the preceding year of the baseline was
assessed by FFQ from ``never'', ``1-2 times/month'', ``1-2 times/week'',
``3-4 times/week'', and ``Almost daily''. The exact amount of milk
consumption was difficult to assess here. However, good reproducibility
and validity were confirmed previously (Spearman rank correlation
coefficient between milk intake frequency and weighed dietary record for
12 days was 0.65) \citep{Date2005}. From the validation study
\citep{Date2005} of the FFQ, daily consumers of milk was found to have a
median intake of 146 gram per day.

>>>>

Moreover, a more intuitive interpretation would be available if we were
able to show the probabilities and how certain the existing data can
provide evidence about whether drinking milk can delay or lower the
hazard of dying from stroke. This would require transformation of the
research questions into ``Do individuals with milk intake habit have
lower hazard of dying from a stroke event compared with people who never
consume milk?'' Another way of asking the same research question would
be ``Can a event of dying from stroke be postponed by having a milk
intake habit?'' The later question is indeed directly comparing the
length of times before events or the speed from the same departure point
(entry of the study) to a fatal stroke event. Accelerated failure time
models (AFT) under Bayesian framework are convenient tools that would
help avoid worrying about the assumptions of repeatability of a large
scale prospective study, and can calculate the exact proportion of
posterior estimates (a hazard ratio) that is lower than 1. AFT model is
helpful in estimating acceleration factors which can be interpreted as
the speed for developing an event of interest. This speed parameter
shows how faster/slower individuals in one exposure group might have an
event compared to others among different exposure groups
\citep{Wei1992, ibrahim2014b}.

Our aim was to provide a more straightforward answer to the primary
research question that whether someone answered he/she drank milk at the
baseline of study had lower hazard of dying from stroke compared with
his/her counterparts who said they never consumed milk. If the answer to
the primary objective was yes, then the probabilities that individuals
with different frequencies of milk intake may had lower hazard compared
with those who never drank milk were calculated through a Markov Chain
Monte Carlo (MCMC) simulation process. A Bayesian survival analysis
method was applied on an existing database and through which, we
provided estimates about whether drinking milk could delay a stroke
mortality event from happening after controlling for the other potential
confounders.

@@

Frequency of milk intake during the preceding year of the baseline was
assessed by FFQ from ``never'', ``1-2 times/month'', ``1-2 times/week'',
``3-4 times/week'', and ``Almost daily''. The exact amount of milk
consumption was difficult to assess here. However, good reproducibility
and validity were confirmed previously (Spearman rank correlation
coefficient between milk intake frequency and weighed dietary record for
12 days was 0.65) \citep{Date2005}. From the same validation study
\citep{Date2005} on the FFQ, daily consumers of milk was found to have a
median intake of 146 gram per day. In the baseline year of the JACC
study (between 1988 and 1990), most ``milk and dairy product''
consumption (92.1\%) was in the form of whole milk \citep{eigyogenjyo}.
